---
title: "pizza_EDA_R"
output:
  #pdf_document: default
  #html_document: default
  output: md_document
date: "2024-05-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the necessary libraries and the dataset
```{r}
library(tidyverse)
library(glue)
# load dataset
df <- read_csv('pizza_sales_clean.csv')
```



## KPI analysis

## Total Revenue

```{r}
glue('The total sales is ${round(sum(df$total_price), 2)}')
```

## Average Order Value

```{r}
glue('The total sales is ${round(mean(df$total_price), 2)}')
```

## Total Pizza Sold

```{r}
glue('The total number of pizzas sold is {sum(df$quantity)}.')
```

## Total Orders

```{r}
glue('A total of {nrow(df)} were placed.')
```
## Average Pizzas Per Order

```{r}
glue('The average number of pizzas ordered by order is {round(mean(df$quantity), 2)}')
```

# Sales Performance Analysis

## What is the average unit price and revenue of pizza across different categories?

```{r}
df %>% group_by(pizza_category) %>% summarise(
	avg_unit_price = round(mean(unit_price),2),
	avg_revenue = round(mean(total_price),2 ) )
```

## What is the average unit price and revenue of pizza across different sizes?

```{r}
# convert the size to a factor variable
df$pizza_size <- factor(df$pizza_size,
						levels = unlist(str_split('S M L XL XXL', ' ')),
						ordered=T)
df %>% group_by(pizza_size) %>% summarise(
	avg_unit_price = round(mean(unit_price),2),
	avg_revenue = round(mean(total_price),2 ) )
```

## What is the average unit price and revenue of most sold 3 pizzas?

```{r}
df %>% group_by(pizza_name) %>% summarise(
	units_sold = sum(quantity),
	avg_unit_price = round(mean(unit_price),2),
	avg_revenue = round(mean(total_price),2 ) ) %>%
	arrange(-units_sold) %>% head(3)
```
# Seasonal Analysis

## Which days of the week have the highest number of orders?

```{r}
library(lubridate)
df %>% group_by(week_day = wday(df$order_timestamp, label=T, abbr = F)) %>% 
	summarise(total_pizzas_ordered = sum(quantity)) %>% arrange(-total_pizzas_ordered)
```

## At what time do most orders occur?

```{r}
df %>% group_by(day_hour = hour(df$order_timestamp)) %>% 
	summarise(total_pizzas_ordered = sum(quantity)) %>% arrange(-total_pizzas_ordered)
```

## Which month has the highest revenue?

```{r}
df %>% group_by(month = month(df$order_timestamp,
							  label=T, abbr=F)) %>% 
	summarise(total_revenue = sum(total_price)) %>% arrange(-total_revenue)
```

## Which season has the highest revenue?

```{r}
# function to classify the month
classifySeason <- function(month) {
  # Ensure the input is in lowercase for uniformity
  month <- tolower(month)
  
  # Use switch statement to determine the season
  season <- switch(month,
                   "january" = "Winter",
                   "february" = "Winter",
                   "march" = "Spring",
                   "april" = "Spring",
                   "may" = "Spring",
                   "june" = "Summer",
                   "july" = "Summer",
                   "august" = "Summer",
                   "september" = "Autumn",
                   "october" = "Autumn",
                   "november" = "Autumn",
                   "december" = "Winter",
                   "Unknown month")  # Default case if the month doesn't match
  return(season)
}
```

```{r}
# create a copy of the dataset
df2 <- df

# get the season
df2$season <- sapply(month(df2$order_timestamp, label=T, abbr=F),
					 FUN=classifySeason)

# get the ranking
df2 %>% group_by(season) %>% summarise(
	total_revenue = sum(total_price) ) %>% arrange(-total_revenue)

# remove the copy of the dataset
rm(df2)
```

# Customer Behavior Analysis

## Which pizza is the favorite of customers (most ordered pizza)?

```{r}
df %>% group_by(pizza_name) %>% summarise(pizzas_ordered = sum(quantity)) %>% arrange(-pizzas_ordered)
```

## Which pizza is ordered the most number of times?

```{r}
most_ordered_pizza <- df %>% group_by(pizza_name) %>% summarise(times_ordered = sum(quantity)) %>% arrange(-times_ordered) %>% head(1)

glue('The most ordered pizza is {most_ordered_pizza$pizza_name}. It was ordered {most_ordered_pizza$times_ordered} times.')
```

## Which pizza size is preferred by customers?

```{r}
sizes_value_counts <- table(df$pizza_size)
most_ordered <- sizes_value_counts[
	which(sizes_value_counts == max(sizes_value_counts)) ]
glue('The preferred pizza size is {names(most_ordered)}, ordered {unname(most_ordered)} times.')
```

## Which pizza category is preferred by customers?

```{r}
categories_value_counts <- table(df$pizza_category)
most_ordered <- categories_value_counts[
	which(categories_value_counts == max(categories_value_counts)) ]
glue('The preferred pizza category is {names(most_ordered)}, ordered {unname(most_ordered)} times.')

```

# Pizza Analysis

## The pizza with the least price and highest price

```{r}
# get list of pizzas and their prices
pizza_names_prices <- unique(df[c("pizza_name", "unit_price")])
# cheapest one
cheapest_pizza = pizza_names_prices %>%
	filter(unit_price == min(pizza_names_prices$unit_price) )
# most expensive one
expensive_pizza = pizza_names_prices %>%
	filter(unit_price == max(pizza_names_prices$unit_price) )
											   
glue('The cheapest pizza is {cheapest_pizza$pizza_name} (${cheapest_pizza$unit_price}).')

glue('The most expensive pizza is {expensive_pizza$pizza_name} (${expensive_pizza$unit_price}).')
```

## Number of pizzas per category

```{r}
df %>% group_by(pizza_category) %>% summarise(
	total_pizzas_ordered = sum(quantity) ) %>%
	arrange(-total_pizzas_ordered)
```

## Number of pizzas per size

```{r}
df %>% group_by(pizza_size) %>% summarise(
	total_pizzas_ordered = sum(quantity) ) %>%
	arrange(-total_pizzas_ordered)
```

## Pizzas with more than one category

```{r}
pizza_more_1_category =any(
	table(
		unique(df[c('pizza_name', 'pizza_category')])
		) > 1
	) 
glue('Are there any pizzas belonging to more than one category? {pizza_more_1_category}')
```

